dyngen: simulating developing single cells

### 2019-09-09 content.tex
convert showcase
fix references

Our simulator draws inspiration from a simulator for bulk transcriptomics data, GeneNetWeaver \cite{schaffter_genenetweaversilicobenchmark_2011,marbach_wisdomcrowdsrobust_2012}, but with two key improvements to make it work for single-cell data. First, instead of simulating continuous systems using stochastic differential equations (SDE), we simulate individual molecules and their reactions using derivatives of Gillespie's Stochastic Simulation Algorithm (SSA) \cite{gillespie_exactstochasticsimulation_1977}. Such a simulation better captures the stochasticity when low number of molecules are present  \cite{gillespie_exactstochasticsimulation_1977}, as is the case in single-cells \cite{zenklusen_singlernacountingreveals_2008a}. Processes such as transcriptional bursting may be difficult to represent in a continuous model, but develop naturally in single-molecule simulations \cite{schuh_genenetworkstranscriptional_2019}. A second improvement is that we include several ways to construct a GRN such that it mimics a dynamic process of interest, such as cell differentiation into multiple cell types.

@@

\begin{figure}[htb!]
	\centering
	\includegraphics[width=\linewidth]{fig/showcase_3.png} 
	\caption{Showcase of dyngen functionality. \textbf{TODO: change to pdf.} Remove B? \textit{Yes please}} % TODO: update label
	\label{fig:dyngen}
\end{figure}

>>>>

Our simulator draws inspiration from a simulator for bulk transcriptomics data, GeneNetWeaver \cite{schaffter_genenetweaversilicobenchmark_2011,marbach_wisdomcrowdsrobust_2012}, but with two key improvements to make it work for single-cell data. First, instead of simulating continuous systems using stochastic differential equations (SDE), we simulate individual molecules and their reactions using derivatives of Gillespie's Stochastic Simulation Algorithm (SSA) \cite{gillespie_exactstochasticsimulation_1977}. Such a simulation better captures the stochasticity when low number of molecules are present  \cite{gillespie_exactstochasticsimulation_1977}, as is the case in single-cells \cite{zenklusen_singlernacountingreveals_2008}. Processes such as transcriptional bursting may be difficult to represent in a continuous model, but develop naturally in single-molecule simulations \cite{schuh_genenetworkstranscriptional_2019}. A second improvement is that we include several ways to construct a GRN such that it mimics a dynamic process of interest, such as cell differentiation into multiple cell types.

@@

\begin{figure}[htb!]
	\centering
	\includegraphics[width=\linewidth]{fig/showcase_4} 
	\caption{Showcase of dyngen functionality. Remove B? \textit{Yes please}} % TODO: update label
	\label{fig:dyngen}
\end{figure}
